\chapter{Creating a Document}
\section{Titular Information}
%

Titular information is declared in the preamble after the package
declaration.  It varies greatly between document classes.  However,
key pieces of information which most documents must have, i.e. title,
author and date; academic documents also require author affiliations,
and some documents require subtitles.  It's important to note that all
these are added at the user's behest.\footnote{Contrary to word
  processors, \LaTeX~will \emph{not} automatically do anything unless
  the user expressly tells it to.} This information is added last in
the preamble, just before the document proper.
\begin{verbatim}
	\title{Document Title}
	\subtitle{Document Subtitle}
	\author{auth_1\thanks{affil_1}, auth_2\thanks{affil_2},...}
	\date{<date>}
\end{verbatim}
The date can be dynamically updated to the OS' by using \verb|\today|
as the \verb|\date{}| command's argument.  The affiliation is added as
a footnote by the \verb|\thanks{}| command.

The preferred way to add author affiliations is by using the
\pkg{authblk} package:
\begin{verbatim}
	\author[x1]{author_1}
	\affil[x1]{affil_1}
	\author[x2]{author_2}
	\affil[x2]{affil_2}
	...
	\author[xn]{author_n}
	\affil[xn]{affil_n}
\end{verbatim}
This package makes keeping track of authors and affiliations much
easier because it relies on labels rather than a command's position.
It also tidies and simplifies having multiple affiliations per author,
and multiple authors per affiliation.

Printing the titular information requires the command
\verb|\maketitle| to be inside the document environment.

%
\section{Paging}
%

Page numbering and formatting are two of the most important, yet most
overlooked things in document typesetting.  There are a few paging
commands and styles but the common ones are found in table
\ref{t:page}.  The \verb|\pagenumbering{<style_name>}| changes the
page numbering to one of the default styles and resets the counter to
its default value of 1.
\begin{table}[!htbp]
    \centering
    \caption{Common paging commands.}
    \label{t:page}
    \begin{tabular}{rl}
        \toprule
        Command & Function \\
        \midrule
        \verb|\pagebreak| & New page with spread out paragraphs. \\
        \verb|\newpage| & New page with normal paragraphs. \\
        \verb|\clearpage| & New page that cannot automatically have floats. \\
        \verb|\cleardoublepage| & Sale as \verb|clearpage| but on two pages. \\
        \bottomrule
    \end{tabular}
\end{table}

There are also paging styles which modify a single page's style.  The
most common default page styles and their function are found in table
\ref{t:pagestyle}.  They are declared within the \verb|\pagestyle{}|
(all pages), or \verb|\thispagestyle{}| (this page only) commands.
\begin{table}[!htbp]
    \centering
    \caption{Common default page styles.}
    \label{t:pagestyle}
    \begin{tabular}{rl}
        \toprule
        Style & Function \\
        \midrule
        \verb|empty| & Blank header and footer. \\
        \verb|plain| & Blank header, footer contains a centered page number.\\
        \verb|headings| & Blank footer, header contains page number \& relevant info. \\
        \bottomrule
    \end{tabular}
\end{table}

%
There are other pace styles defined in packages such as
\pkg{fancyhdr}, whose manuals and downloads can be found on the net.

\subsection{Page Format}
%

The standard \LaTeX~page formatting is usually sufficient for everyday
use.  However there are times---such as when writing a thesis template
and short but comprehensive \LaTeX~tutorial---when you need special
formatting options.  Such options follow the structure:
\begin{itemize}
    \item \verb|\setlength{\lengthcommand}{<len>}|: set a length command to a new value \verb|<len>|.
    \item \verb|\addtolength{\lengthcommand}{<len>}|: add a value, \verb|<len>|, to a length parameter.
    \item \verb|\linespread{<len>}|: set inter-line spacing to the value \verb|<len>|.
\end{itemize}
Added lengths can be negative, so it is possible to subtract lengths.
It's also possible to use factors of one length to define another by
placing the factor before a specific \verb|\lengthcommand|,
e.g. \verb|-0.5\textwidth| divides the text width by $-2$.

There are \verb|\lengthcommand|s for all manner of things, from margin
padding to paragraph skip length---and the need to modify them is
niche at best, so most are excluded from this guide.  More information
on lengths can be found in the lengths entry of the
\LaTeX~wikibook\footnote{\url{https://en.wikibooks.org/wiki/LaTeX/Lengths}}.
The most used by far are \verb|\textwidth|---text width, which is
often different from the page width due to text padding and
multicolumns---and \verb|\width|---page width with margins---because
of how useful they are in scaling images.

%
\section{Document Environment}
%
%

The \verb|document| environment is where the document's actual content
is found.  It is the master environment within which all other
environments must be called.  Thus the first line after the preamble
is \emph{always} \verb|\begin{document}| and the \texttt{.tex} file's
very last line is \emph{always} \verb|\end{document}|.

\section{Document Sectioning}
%

A standard document can be sectioned into parts, chapters, sections,
subsections and subsubsections.  However, the user can add new and
edit existing sectioning commands to change their format, or even add
more arguments.  It is quite well known---in \emph{most} reputable
moustache-twirling circles---that should you require more sectioning
levels than the \LaTeX~default, you are most likely doing it
\emph{wrong}, and will forever be cursed with spaghetti code.  Don't
let it happen to you, structure your thoughts, documents, and code
properly.  The standard sectioning commands are declared as follows:
\begin{itemize}
	\item \verb|\part{<part_name>}|: Parts. Primarily used in book writing.
	\item \verb|\chapter{<chapter_title>}|: Chapters.
	\item \verb|\section{<section_title>}|: Sections.
	\item \verb|\subsection{<subsection_title}>|: Subsections.
	\item \verb|\subsubsection{<subsubsection_title>}|: Subsubsections.
\end{itemize}
%
Their starred versions prevents them from being numbered, and stops
them from being indexed unless they are manually added to the table of
contents (see section \ref{s:contents}).
